---
title: "In-class_Ex04"
date: "6 May 2023"
today_date: "'r Sys.Date()'"
execute: 
  warning: false
  echo: true
  eval: true
editor: visual
---

# Install Packages

-   rstatix: perform statistic tests and store the result in tabulate form instead of list conventionally

-   gt: create distribution statistic table

-   patchwork: combine multiple plots into a single plot or arrange them in a grid layout

```{r}
pacman::p_load(rstatix, gt, patchwork, tidyverse)
```

# Read Data

```{r}
exam_data<- read_csv('data/Exam_data.csv')
```

# QQplot

If the data is normally distributed, the points will lie on a straight diagonal red line.

::: panel-tabset
## The plot

```{r}
#| echo: false
#| fig-width: 4
#| fig-height: 4
ggplot(exam_data,
       aes(sample=ENGLISH))+
  stat_qq()+
  stat_qq_line(col='red')
```

## The code

```{r}
#| eval: false
#| fig-width: 4
#| fig-height: 4
ggplot(exam_data,
       aes(sample=ENGLISH))+
  stat_qq()+
  stat_qq_line(col='red')
```
:::

::: callout-note
we can see that the points deviate significantly from the straight diagonal line. This is a clear indication that the set of data is not normally distributed.
:::

::: panel-tabset
## The plot

```{r}
#| echo: false
qq<- ggplot(exam_data,
            aes(sample=ENGLISH))+
  stat_qq()+
  stat_qq_line()

sw_t<- exam_data %>%
  shapiro_test(ENGLISH) %>%
  gt()

tmp<- tempfile(fileext = '.png')
gtsave(sw_t, tmp)
table_png<- png::readPNG(tmp, native = TRUE)

qq+table_png
```

## The code

```{r}
#| eval: false
qq<- ggplot(exam_data,
            aes(sample=ENGLISH))+
  stat_qq()+
  stat_qq_line()

sw_t<- exam_data %>%
  shapiro_test(ENGLISH) %>%
  gt()

tmp <- tempfile(fileext = '.png')
gtsave(sw_t, tmp)
table_png <- png::readPNG(tmp, native = TRUE)

qq+table_png
```
:::
